---
title: "Data Science HW 1"
author: "Riya Bhilegaonkar"
date: "2022-09-17"
output: github_document
---

Load the tidyverse library
```{r}
library(tidyverse)
```

## Problem 1

Load the penguins dataset:

```{r}
data("penguins", package = "palmerpenguins")
```

Descriptions of the `penguins` dataset:
The data in this dataset contains information about different penguins including their species, location (islands of Palmer Archipelago, Antartica (Biscoe, Dream or Torgersen)), bill length (in mm), bill depth (in mm), flipper length (in mm), body mass (in g), sex and the year.

The size of the dataset is 344 x 8.
The number of rows in the dataset is 344.
```{r}
nrow(penguins)
```

The number of columns in the dataset is 8.
```{r}
ncol(penguins)
```

The mean flipper length is 200.9 mm, which we find using the summary function.
```{r}
summary(penguins)
```

Making a scatterplot of `flipper_length_mm` vs `bill_length_mm`
```{r}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()
```

Exporting the scatterplot to project directory using ggsave:
```{r}
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

## Problem 2
Create a data frame comprised of:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r}
p2_df <- tibble(
  norm_sample = rnorm(10),
  norm_samp_pos = norm_sample > 0,
  char_vec = c("a1", "a2", "a3", "a4", "a5", "a6", "a7", "a8", "a9", "a10"),
  factor_vec = factor(c("small", "medium", "large", "small", "medium", "large", "small", "medium", "large", "small"))
)

```

Taking the mean of each variable in the dataframe.
Taking the mean of the `norm_sample` variable:
```{r}
mean(pull(p2_df, norm_sample))
```

The mean of `norm_sample` is 0.1389.

Taking the mean of the `norm_samp_pos` variable:
```{r}
mean(pull(p2_df, norm_samp_pos))
```

The mean of `norm_samp_pos` is 0.7.

Taking the mean of the `char_vec` variable:
```{r, eval = FALSE}
mean(pull(p2_df, char_vec))
```

As the values in the variable are not numeric or logical, the mean does not work and the returns the value of NA. 

Taking the mean of the `factor_vec` variable:
```{r, eval = FALSE}
mean(pull(p2_df, factor_vec))
```

As the values in the variable are not numeric or logical, the mean does not work and the returns the value of NA. 

Now applying the `as.numeric` function 
```{r, eval = FALSE}
mean(pull(p2_df, as.numeric(norm_samp_pos)))
mean(pull(p2_df, as.numeric(char_vec)))
mean(pull(p2_df, as.numeric(factor_vec)))
```

When we use the as.numeric function we similarly obtain NA for the mean of the character vector and the factor vector, as R is not able to convert these multi-level variables or numerous options into true or false values or 0 or 1 as bianry variables. However in the case of the logical vector, R treats TRUE or FALSE as the value of 1 or 0 and hence calculates the mean using the averages of those values. This does help us understand why we get NA values for the means of character or factor vectors versus the case of logical vectors.
